---
title: "Bob Ross Paintings"
author: "Sam Taylor"
date: "14/08/2019"
output: html_document
---

```{r markdown_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Joy of Painting with Bob Ross

This week's data is from 538, and can either be read in directly from this repo or from the 538 R package which was recently updated to include more data from 538!

The 538 article can be found here: https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/ 


```{r setup}
# load packages
library(tidyverse)
library(janitor)
library(tidytext)
library(topicmodels)
library(wordcloud)

# set theme
theme_set(theme_bw())

# load data
bob_ross <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-06/bob-ross.csv")

# tidy data
bob_ross_processed <- bob_ross %>% 
                        janitor::clean_names() %>% 
                        mutate(season = str_sub(episode,2,3),
                               title = str_replace_all(str_to_lower(title),'"', ""))
```

## What are the most common works of Bob Ross?

By far the most common paintings which Bob produces are of trees. Clouds and mountains also
feature prominantly in his works. 

```{r common_subjects}
bob_ross_processed %>% 
  gather(key, value, -season, -title, -episode) %>% 
  group_by(key) %>%
  summarise(total_paintings = sum(value)) %>% 
  top_n(16) %>% 
  ggplot(aes(fct_reorder(str_to_title(key),total_paintings), total_paintings,
             fill=total_paintings)) + 
    geom_col(show.legend=F, color='black') + 
    coord_flip() +
    labs(x='',
         y='Total Paintings',
         title = 'Trees are the most common subject of Bob Ross paintings') +
    scale_fill_gradient2(high='dodgerblue2', midpoint = 134)
```

## Have the works of Bob Ross been changing over time?

Although some themes are constant in Bob Ross paintings, some have changed over time.
 
```{r themes}
# transform to long
bob_ross_long <- bob_ross_processed %>% 
  select(-title, -episode) %>% 
  gather(key, value, -season)
# get top 10 subjects
top_subjects <- bob_ross_processed %>% 
  select(-title, -episode,-season) %>% 
  gather(key, value) %>% 
  group_by(key) %>%  
  summarise(season_paintings = sum(value),
            total_paintings = n(), 
            perc_total = season_paintings/total_paintings) %>% 
  top_n(6) %>% 
  select(key)
# plot the change over time
bob_ross_long %>% 
  filter(key %in% top_subjects$key) %>% 
  group_by(season, key) %>% 
  summarise(total_paintings = sum(value)) %>% 
  ggplot(aes(season, total_paintings, fill=str_to_title(key))) + geom_col(position='fill') +
  labs(x='Season',
       y='Proportion of Paintings',
       fill='Painting Subject',
       title = 'Clouds have become less prominant a theme in Bob Ross paintings over time') 

```
 
## What are the most titles of Bob's work 

Given the subject matter of the paintings it's not surprising to see common words in the
title's of the paintings reflect this.

```{r wordcloud}
bob_ross_processed %>% 
  select(season, title) %>% 
  unnest_tokens(word, title) %>% 
  anti_join(stop_words) %>% 
  count(word, sort=T) %>%
  with(wordcloud(word,n, max.words=150)) 
```
 
 