---
title: "Meteorite Impacts"
author: "Sam Taylor"
date: "11/06/2019"
output: html_document
---

This week's dataset is a dataset all about meteorites, where they fell and when they fell! Data comes from the Meteoritical Society by way of NASA. H/t to #TidyTuesday community member Malin Axelsson for sharing this data as an issue on GitHub!

https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-11 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}

# load packages
library(tidyverse)
library(gganimate)
library(magick)

# set default ggplot theme
theme_set(theme_bw())

# read in data
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv") 

# tidy data as appropriate - removing id 32789 as appears to be bad data.
meteorites_processed <- meteorites %>% 
                          mutate(decade = 10 * (year %/% 10),
                                 class = fct_lump(class, n=8)) %>%
                          filter(id != 32789) 

```

Recording of meteorite impacts hav become more common over time

```{r, warning=FALSE, message=FALSE}

# removed meteorites period to 1950 as generally aren't really recorded. In addition, anything post 2019 (current year) is also removed.

meteorites_processed %>% 
  filter(year >1950 & year < 2019) %>% 
  ggplot(aes(x=year)) + geom_histogram(binwidth = 1) +
  labs(y="", x='Decade of meteorite fall', title="Recording of meteorite falls only really began post 1970")

```

Mass of meteorites are log normally distributed. 

```{r, warning=FALSE, message=FALSE}

meteorites_processed %>% 
  ggplot(aes(log(mass))) + geom_histogram() +
  labs(y='',
       x='Meteorite Mass (Log Scale)',
       title ='Mass of meteorites is log normally distributed')
```

Meteorites have gradually getting large since the 1970s although not as large prior to this point in time. NB - the sample size of falls per decade is significantly smaller for 1950s and 1960s giving much more uncertainty around the estimates.

```{r, warning=FALSE, message=FALSE}
meteorites_processed %>% 
  filter(year >1950 & year < 2019) %>% 
  ggplot(aes(x=decade, y=log10(mass), group=decade)) + geom_boxplot(notch=T) +
  labs(x="Decade",y="Mass of Meteorites (Log2 Scale)",
       title="Meteorites having been getting heavier over time since the 1970s") 
```

When regressed against decade, the size of meteorite appears to reduce by 2808.8g per decade.

```{r, warning=FALSE, message=FALSE}
lm(mass~decade, data=meteorites_processed) %>% 
  summary()
```

The plot below shows changing volumes of meteorite impacts and their associated weight since 1950s. Primary changes appears to be actually recording strikes in Africa in circa 1990 onwards.

```{r, warning=FALSE, message=FALSE}

meteorites_processed %>% 
  filter(year >1950 & year < 2019) %>% 
  ggplot(aes(long, lat)) +    
  geom_point(aes(color=log(mass)),alpha=1/8) +
  borders(database = "world") + theme_void() + 
  scale_color_continuous(low='light blue',
                         high='navy blue',
                         name ='Meteorite \n Mass') +
  theme(legend.position = c(0.1, 0.4),
        plot.title = element_text(size=25, face='bold',
                                  hjust = 0.5, vjust=0),
        plot.caption = element_text(size=14)) +
  transition_manual(frames=decade) +
  enter_fade() + 
  exit_shrink() +
  coord_map(xlim=c(-180,180),
            ylim=c(-60, 90)) + 
  ease_aes('sine-in-out') +
  labs(title="Meteorite Impacts since the 1950s",
       subtitle = "",
       caption = 'Decade: {current_frame}')

```

