---
title: "UFO"
author: "Sam Taylor"
date: "25/06/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
# load packages
library(tidyverse)
library(lubridate)
library(countrycode)
library(gganimate)

# read in data.
ufo_sightings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")

View(ufo_sightings)

# tidy data for mapping - tidy dates and city area strings
ufo_sightings_processed <- ufo_sightings %>% 
                            mutate(date_time = mdy_hm(date_time),
                                   year = floor_date(date_time, unit = 'year'),
                                   city_area = str_to_lower(str_remove(city_area, '\\((.*?)\\)')),
                                   ufo_shape = ifelse(ufo_shape =='unknown' | is.na(ufo_shape), 'other', ufo_shape)) %>%
                            mutate(ufo_shape = fct_lump(str_to_title(ufo_shape),5))
                            dplyr::filter(!is.na(country))
                            
# add in the country code  
ufo_sightings_processed$country <- countrycode(sourcevar =ufo_sightings_processed$country, origin='iso2c',destination='country.name')


View(ufo_sightings_processed)

```

## How have reports of UFO sightings changed over time?

UFO sightings were rarely recorded prior to 1920 with the majority of records 
being missing data prior to this point (Although in all cases bar one, the 
specific location of the incident is recorded). The most commonly recorded sightings 
are in North America, particually in the USA.

```{r}
ufo_sightings_processed %>% 
  group_by(country, year) %>% 
  dplyr::count() %>% 
  ggplot(aes(x=year, y=log(n), color=country)) + geom_line() 

```

## Where is the best place to spot a UFO?

Your best chance at spotting a UFO is likely to be in America - all about one of the most common city areas are there! 

```{r}
ufo_sightings_processed %>% 
  dplyr::count(city=str_to_title(city_area), sort=T) %>% 
  top_n(10) %>% 
  ggplot(aes(x=fct_reorder(city,n), fill=city, y=n)) + 
  geom_col(show.legend = F) + coord_flip() +
  labs(x='', y='Number of reported UFO sightings',
       title ='Which cities have the most reported UFO sightings?') 
```

```{r}

test <-ufo_sightings_processed %>% 
  ggplot(aes(longitude, latitude, shape=ufo_shape, color=log(encounter_length/60))) + 
    geom_point(alpha=1/3) +
    coord_map(xlim=c(-180,180),
              ylim=c(-60, 90)) + 
    borders(database = "world") + 
    theme_void() +
    scale_color_continuous(low='light blue',
                           high='navy blue',
                           name ='Encounter Length \n (Log Minutes)') +
    scale_shape_discrete(name='UFO Shape') +
    theme(legend.text = element_text(size=6),
          legend.title = element_text(size=7),
          plot.title = element_text(size=16, face='bold',
                                  hjust = 0.1, vjust=0),
          plot.caption = element_text(size=10)) +
    labs(title="UFO Sights since the 1920s",
         subtitle = "",
         caption = 'Year: {current_frame}') +
    transition_manual(frames=str_sub(year,1,4)) +
    enter_fade() + 
    exit_shrink() +
    ease_aes('sine-in-out') 

anim_save("amin.gif", animation = test)

```

do something with text
have ufo encounters been getting long over time?
have the types of encounters been changing?


