---
title: "pizza_reviews"
author: "Sam Taylor"
date: "07/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pizza Reviews

This week's data is from Jared Lander and Barstool Sports via Tyler Richards.

Credit for this week's concept goes to Ludmila who did a recent dataviz presentation and gave shoutouts to both #tidytuesday and a pizza dataset!

Check out her DataViz video and slides at her GitHub

Jared's data is from top NY pizza restaurants, with a 6 point likert scale survey on ratings. The Barstool sports dataset has critic, public, and the Barstool Staff's rating as well as pricing, location, and geo-location. There are 22 pizza places that overlap between the two datasets.

If you want to look more at geo-location of pizza places, checkout this one from DataFiniti. This includes 10000 pizza places, their price ranges and geo-locations.

```{r set_up, echo=FALSE, messages=FALSE}
# Load packages
library(tidyverse)
library(ggthemes)
library(lintr)

# Load data
pizza_jared <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_jared.csv")
pizza_barstool <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_barstool.csv")
pizza_datafiniti <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_datafiniti.csv")

```

## Preprocessing

Data Infiniti appears to have duplicate data and although it initially appears to have 10,000 data points. It only actually has `r pizza_datafiniti %>% n_distinct()`. These will be removed for the purposes of this analysis and only distinct locations will be included.

```{r preprocessing}
# Remove duplicate values for the dataset
pizza_df <- pizza_datafiniti %>% 
  distinct() %>% 
  select(-country)

```

## EDA


```{r}
# Which states have the most pizza reviews?

pizza_df %>% 
  count(province, sort=T) %>% 
  top_n(10) %>% 
  ggplot(aes(reorder(province,n), n, label=n)) + 
    geom_col() +
    geom_text(hjust=-0.25) +
    labs(x='',
         y='',
         title='Which states have the most pizza reviews?') +
    theme_fivethirtyeight() + coord_flip()

```





```

